---
title: "Section 2"
author: "Lily, James, Tobin"
date: "30/05/2018"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(GGally)
source('conf_int_prop.R')
```

# Part B: Logistic Regression

## Introduction 
Breast cancer was the second most commonly diagnosed cancer in Australia in 2013 [^1]. The most effecitive method for breast cancer screening is Mammography. To confirm the diagnoisis, invasive biopsies are perfomed. However, 70% of biopsies come back benign, indicating a high false positive rate in Mammography. To improve this process several computer-aided diagnosis (CAD) systems have been developed to help aid clinicians in making informed diagnoses.

The data in this examination contains 961 mammographic mass lesions with 445 of those lesions being malignant, given by the indicator varible Severity. Addionally for each of these lesions there are three attributes from the Breast Imaging Reporting and Data System (BI-RADS), including the lesion shape (round= 1, oval= 2, lobular= 3, irregular= 4), the margin (circumscribed= 1, microlobulated= 2, obscured= 3, ill-defined= 4, spiculated= 5) and the density (high= 1, iso= 2, low= 3, fat-containing= 4). 

This apprach will attempt to develop a logistic predictive model for mammographic mass severity using the available predictor variables and to obtain predicted probabilities of mass severity that can used by clinicians to make informed diagnoses.


[^1]: Australian Institute of Health and Welfare, Cancer compendium: information and trends by cancer type, https://www.aihw.gov.au/reports/cancer/cancer-compendium-information-and-trends-by-cancer-type/report-contents/breast-cancer-in-australia,  [Accessed May 2018].

### Data cleaning


```{r}
mammo <- read.csv('mammo.txt', header = TRUE)
head(mammo)
```

After taking the data it is clear there are a few key issues to deal with in cleaning the data. The first is the incorrect classes of several of the varibles and, as can be seen below, there are a number of data points that are missing certain attributes, these are currently set to "?". but should be set to "NA", for approciate use within the models. These data points could be removed as they are missing some data, however this should not be done for two reasons. Firstly, incomplete data should not be thrown away as it may still contain valueble ifnormation. Furturemore, the final model may not include some of the predicor attributes, and some of the currently incomplete data may not be missing infomartion in any of the include attributes. 

```{r}
table(mammo$Severity)
```
Severity has no missing data.
```{r}
table(mammo$Age) 
```
Age has 5 missing data points.
```{r}
table(mammo$Shape) 
```
Shape has 31 missing data points.

```{r}
table(mammo$Margin)
```
Margin has 48 missing data points.

```{r}
table(mammo$Density) 
```
Density has 76 missing data points.

```{r}
table(mammo$BI.RADS)
```
BI-RADS has 2 missing data points, but this class will not be used in our model, and is not of high importance.


This data is now cleaned by setting "?"'s to NA's and fixing the attribute classes.
```{r}
class(mammo$Age)
mammo$Age[mammo$Age == "?"] <- NA
mammo$Age <- as.numeric(mammo$Age)
summary(mammo$Age, exclude = FALSE) 
```
Age has its 5 missing data points set to NA and the varible is set to numeric. 
```{r}
class(mammo$Shape)
mammo$Shape <- as.character(mammo$Shape)
mammo$Shape[mammo$Shape == "?"] <- NA
mammo$Shape <- factor(mammo$Shape)
summary(mammo$Shape, exclude = FALSE) 
```
Shape has its 31 missing data points set to NA and the varible is set to a factor 

```{r}
class(mammo$Margin)
mammo$Margin <- as.character(mammo$Margin)
mammo$Margin[mammo$Margin == "?"] <- NA
mammo$Margin <- factor(mammo$Margin)
summary(mammo$Margin, exclude = FALSE) 
```
Margin has its 48 missing data points set to NA and the varible is set to a factor 

```{r}
class(mammo$Density)
mammo$Density <- as.character(mammo$Density)
mammo$Density[mammo$Density == "?"] <- NA
mammo$Density <- factor(mammo$Density)
summary(mammo$Density, exclude = FALSE) 
```
Density has its 76 missing data points set to NA and the varible is set to a factor 

```{r}
class(mammo$Severity)
mammo$Severity <- as.numeric(mammo$Severity)
summary(mammo$Severity, exclude = FALSE)
```
Severity is set to a integer.


Finally, BI-RADS has it's 2 missing data points set to NA. Addintioally, the data has a clear outlier in it that is set to NA aswell. Again, this is not overly important as BI-RADS will not be used as a predictor in this model.
```{r}
class(mammo$BI.RADS)
mammo$BI.RADS <- as.character(mammo$BI.RADS)
mammo$BI.RADS[mammo$BI.RADS == "?"] <- NA
mammo$BI.RADS <- factor(mammo$BI.RADS)
summary(mammo$BI.RADS, exclude = FALSE) # 2 NAs, 1 outlier
mammo$BI.RADS[mammo$BI.RADS == 55] <- NA # Set outlier to NA
mammo$BI.RADS <- as.numeric(mammo$BI.RADS) 
summary(mammo$BI.RADS, exclude = FALSE) # 3 NAs
```

# TODO: Ask if Age is forced to be a parameter (it ends up in there anyway)
# Can these varibles be interpolated (no)

## Data Visualisation
To examine the relationship between each of the indivual varibles, a plot of the relationships between the varible is made. 
```{r, warning=FALSE}
mammo %>%
  mutate(Severity = as.factor(Severity)) %>%
  select(2:6) %>%
  ggpairs(upper = list(continuous = "blank",
                       combo ="blank",
                       discrete = "blank",
                       na = "blank"),
          lower = list(continuous = "cor",
                       combo = "box_no_facet",
                       discrete = "facetbar",
                       na = "na")) 
```

Now I'll talk about the relationships between varibles but I'm sleepy and frankly that density is giving me the middle finger so I'll do it tomorrow.

Anyway,

## Making a Model

### Using the right data
Firstly, we are only able to fit a model to entries in the data frame that are complete. To do this, we created a logical vector for all the data that references which data entries have information for all variables.

``` {r complete}
complete=!is.na(mammo$Age)&!is.na(mammo$Shape)&!is.na(mammo$Margin)&!is.na(mammo$Density) # Use complete data for models
```

### The Full Model

As we begin the model creation process, our first step is to create a model that uses all of our predictor variables to model our response variable, severity. Of the 4 predictor variables, only age is a numerical variable. Shape, Margin, and Density are all categorical variables which contain levels. In r, when fitting a model to categorical variables, the model uses a reference category. The reference category for Shape should clearly be the combined 1 and 2 level, as each level seems to show a distinct difference in the proportion of malignant masses, and this combined level contains the greatest number of data entries. However for margin, choosing a reference category because more difficult. The default reference level 1 is not the best selection. As we saw from the confidence intervals, category 1 has a relatively low proportion malignant mass cases, while the other categories 2,3,4, and 5 do not deviate from one another all that much. With 1 as the reference level, all the other levels appear significantly different, when really it is 1 that is the oddball. Hence we must decide on another reference level. Category 4 has the largest number of data points, and the true proportion of malignant cases with a margin of 4 has a narrow range of possibilities relative to the other categories, meaning any other facets of the data will be more prevalent when fitting the model, therefore this is the reference point we will use.

``` {r full}
mammo$Margin <- relevel(mammo$Margin, ref="1")

mod.full1 <- glm(formula = Severity ~ Age + Shape + Margin + Density,
               family = binomial,
               data = mammo[complete,])

table(mammo$Margin)

mammo$Margin <- relevel(mammo$Margin, ref="4")

mod.full4 <- glm(formula = Severity ~ Age + Shape + Margin + Density,
                 family = binomial,
                 data = mammo[complete,])


summary(mod.full1)
summary(mod.full4)

mod.full <- mod.full4
```

### Model by stepwise selection

Pretty self explanatory: find the scale, r fits the model based on AIC values.

``` {r stepwise}
s2 <- sum((mod.full$residuals)^2)/mod.full$df.residual
s2

# Stepwise Selection

mod.step <- step(mod.full, scale = s2)
summary(mod.step)
```



```{r compare}
summary(mod.full)
summary(mod.back)
summary(mod.full)
qchisq(1-0.05, df = 7)
mod.step$deviance - mod.full$deviance 
anova(mod.step, mod.full) # reject hypothesis that all coefficients are equal

qchisq(1-0.05, df = 2)
mod.step$deviance - mod.back$deviance
anova(mod.step, mod.back) # These literally say the same thing but in the prac we use the difference

AIC(mod.full, mod.step, mod.back)
```
